State of the Practice for Lattice Boltzmann Method Software

### 2022-01-13 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to add information on comparison to other research software

@@

@@

A lack of software development experience was noted by the developer of TCLB.
Many of the team members on their project are domain experts and there can be a
steep learning curve before these members contribute good quality source code.
This problem has been mitigated by re-writing the code to more easily enable
future contributions.  On a related point, improving the software engineering
education or experience of developers is also an idea that was brought up by
several developers.  \wss{The Wiese paper and others talk about the background
of developers.}

@@

Difficulties with ensuring correctness were also noted by several developers.
They indicated that tests are run on all new source code additions, testing both
individual modules and the system to verify correctness. These tests compare the
package output to known correct output using test cases. The developer of TCLB
commented that the amount of testing data that is needed for some cases is a
problem as free testing services do not offer capabilities to store and process
such large amounts of data, and in-house testing solutions needed to be created
to address this limitation. The solution for this has been to limit the size of
the testing problems, and to run tests in small batches with few iterations.
\wss{The Weiss paper alludes this this as well.  I think the Kelly and Sanders
paper likely talks about testing problems and correctness.}

@@

>>>>

Pain points that others, but not LBM developers, mentioned in surveys include
\citep{PintoEtAl2018}: Cross-platform compatibility, interruptions while coding,
scope bloat, lack of user feedback, mismatch between coding skills and
subject-matter skills (for instance, ``I am a mathematician, not an expert on
software''), lack of formal reward system, hard to collaborate on software
projects, aloneness.

@@

Lack of time was also mentioned by \citet{PintoEtAl2018} and
\citet{PintoEtAl2016}.

@@

A lack of software development experience was noted by the developer of TCLB.
Many of the team members on their project are domain experts and there can be a
steep learning curve before these members contribute good quality source code.
This problem has been mitigated by re-writing the code to more easily enable
future contributions.  On a related point, improving the software engineering
education or experience of developers is also an idea that was brought up by
several developers.  \wss{The Wiese paper and others talk about the background
of developers. The knowledge for developing research software primarily comes
from peers and through self-study, not from formal training
\citep{HannayEtAl2009}.  The thinking of the developer on the role of formal education matches the trend \citet{PintoEtAl2018} observes, where in their replication of a previous study \citept{HannayEtAl2009}, they show a growing interest in formal training (From 13\% of respondants in 2009 to 22\% in 2018).}

Poor software documentation is also mentioned by \citet{PintoEtAl2018} and
\citet{LethbridgeEtAl2003}.  The survey from \citet{LethbridgeEtAl2003} is for
software in general, not just research software.

@@

Difficulties with ensuring correctness were also noted by several developers.
They indicated that tests are run on all new source code additions, testing both
individual modules and the system to verify correctness. These tests compare the
package output to known correct output using test cases. The developer of TCLB
commented that the amount of testing data that is needed for some cases is a
problem as free testing services do not offer capabilities to store and process
such large amounts of data, and in-house testing solutions needed to be created
to address this limitation. The solution for this has been to limit the size of
the testing problems, and to run tests in small batches with few iterations.
\wss{The Weiss paper alludes this this as well.  I think the Kelly and Sanders
paper likely talks about testing problems and correctness.  ``while many
scientists believe that software testing is important, a smaller number believe
they have sufficient understanding about testing concepts''
\citep{HannayEtAl2009}.}

@@

``unique challenges of scientific research (e.g., the frequent and unforeseen
changes in requirements ...)'' \citep{PintoEtAl2018}. We can point out that in
the future this viewpoint could change and why.  Not really a unique challenge.
Cite Smith 2016?
